### 2017-01-04 Chapters/experiments.tex
Layout fix, changing size of some figures

\chapter{Dynamics of the gene expression machinery during an acetate-glucose upshift}
\chaptermark{Gene expr. machinery during an upshift}
\label{chap:experiments}

@@

Finalement, nous appliquons cette méthode à la reconstruction du taux de croissance puis de $\alpha (\cdot)$, et présentons les résultats dans la Section~\ref{sec:estim_gr_ra}.
En moyenne, le taux de croissance présente le comportement attendu (Figs~\ref{fig:growth_rate_estimation} et \ref{fig:growth_rate_estimation_median}), avec des valeurs à l'état de croissance stationnaire qui sont proches de celles connues pour \textit{E.~coli}, et une augmentation rapide au moment de la transition vers le milieu riche.
Les résultats pour la reconstruction de $\alpha (\cdot)$ sont cependant plus mitigés.
Même si un comportement oscillatoire est observé pour chaque cellule (Figs~\ref{fig:gene_activity} et \ref{fig:gene_activity_heatmap}), il n'est pas certain que ce dernier ne soit pas un artefact dû à une mauvaise calibration de la méthode.
Il aurait fallu en effet obtenir de longs états stationnaires de croissance avant et après la transition pour pouvoir réaliser cette calibration dans de bonnes conditions.
Néanmoins, si l'on s'intéresse au comportement médian des cellules (Fig.~\ref{fig:gene_activity_median}), on constate que la première oscillation survenant après la transition semble conservée dans la population (Fig.~\ref{fig:gene_activity_heatmap}).
Ces résultats s'avèrent finalement encourageants, et incitent à pousser plus en avant cette étude, notamment via l'acquisition de nouvelles données pour lesquelles les problèmes cités auront pu être corrigés (voir discussion en Section~\ref{sec:chap3_discussion}).

\end{chapter_summary}
\selectlanguage{english}

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/data_acquisition}
\caption{
\textbf{Results of data acquisition.}
We imaged 6 fields (X[1-3]Y[1-2]) each containing 15 wells (W[0-14]) for $\sim$40~h.
From the 90 wells, 68 were suitable for further analysis (the others being empty, out of frame for some period of time, or plugged).
For all wells, data points are missing in the interval between [-720,-150] because the camera was out of focus.
We stopped the analysis at 550~min, when about 3/4 of the bacteria were still growing, before the entire population died within a few hours for an unknown reason.
The image labeled "Raw image" is the last image analyzed for the highlighted well.
The bacterium on the left of this image was manually segmented by selecting two pixels at the poles on the fluorescence images (red cross).
A 6-pixel-wide rectangular mask was computed for each image, resulting in the "Segmented image" on the right.
Fluorescence intensities are expressed in Relative Fluorescence Units (RFU) on a 16-bit image and were corrected for camera background, but not autofluorescence background (Material and Methods~\ref{sec:cell_segmentation} and Discussion~\ref{sec:chap3_discussion}).
The fluorescence intensity of the cell, expressed in units RFU/pixel/cell, was computed by dividing the sum of the fluorescence intensities of the pixels in the rectangle by the total number of pixels in the rectangle.
The cell length is the distance in pixels between the two poles (red line).
The thick lines in green and black highlight the time-varying length and fluorescence density for the cell visible in the top images, labeled "X3Y2, W2".
The vertical dashed lines represent the time of the upshift from growth on acetate to growth on glucose.
}
\label{fig:data_acquisition}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/synthetic_upshift}
\caption{
\textbf{Performance of the Kalman smoothing procedure on synthetic data simulating an acetate-glucose upshift.}
\textit{(A)}~Synthetic data simulating an upshift from acetate to glucose, with and without additive white noise, as well as the results obtained by Kalman smoothing.
The synthetic data were generated by simulating the model presented in Eqs~\ref{eq:pdef}-\ref{eq:rdef} with the on-off regulatory strategy (Eq.~\ref{eq:stratswitch} and Fig.~\ref{fig:onoffresults}).
The model parameters used for the simulation are $e_{M,\texttt{Ace}} = 0.18$~h\textsuperscript{-1}, $e_{M,\texttt{Glu}} = 0.9$~h\textsuperscript{-1}, $k_R = 3.6$~h\textsuperscript{-1}, $\beta = 0.003$~L~g\textsuperscript{-1}, $K_R = 1$~g~L\textsuperscript{-1}.
The predicted $r(t)$ profile was multiplied by a factor $\gamma = 0.02$~RFU~L~g\textsuperscript{-1} in order to obtain the corresponding fluorescence intensity profile $F$ (dashed black curve).
The noise level was estimated from the data (\ref{S7_Text}) and added to $F$.
The choice of the parameters of the Kalman smoothing procedure is discussed in the Material and Methods~\ref{sec:meth_kalman}.
\textit{(B)}~Estimation of the resource allocation profile $\gamma \alpha / \beta$ based on the data in \textit{(A)}.
Following Chapter~\ref{chap:theory}, $\alpha (t)$ displays a bang-bang-singular profile during the upshift (dashed black curve).
While the Kalman smoother is not able to capture the discontinuous variations in $\gamma \alpha / \beta$, it qualitatively reproduces the input quite well (red solid curve).
\textit{(C)}~The predicted growth rate during the upshift experiment.
This information is used as an input in the smoothing procedure, since it is supposed to have been independently estimated from the measurements $\left\{L(t_0), ..., L(t_{N-1}) \right\}$.
}
\label{fig:synthetic_upshift}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/suppinfo_caract}
\caption{
\textbf{Growth curves in M9~0.2\% glucose for the \textit{rpsB-gfp} and \textit{rpsB-mCherry} strains.}
Growth and measurements were performed in a 96-well microplate placed in a Tecan infinite 200 pro reader thermostated at 37$^\circ$C.
We monitored 4 wells inoculated with the WT strain, the \textit{rpsB-gfp} strain, and the \textit{rpsB-mCherry} strain, for a total of 12 wells each containing 150~$\mu$L of growth medium.
Curves were shifted in time to correct for pipetting variability in the inoculation process (-50 min for WT, and -20 min for \textit{rpsB-mCherry}).
\textit{(A)}~The light absorbance as measured by the Tecan microplate reader, corrected for background by removing M9 absorbance.
The superposition of the curves indicates similar growth rates between the WT and the modified strains.
\textit{(B)}~The fluorescence measured in the wells, expressed in Relative Fluorescence Units (RFU).
Green fluorescence (485~nm excitation, 535~nm emission) is measured for the WT (dashed lines) and the \textit{rpsB-gfp} (solid lines) strains.
Red fluorescence (560~nm excitation, 635~nm emission) is measured for the WT (dashed lines) and the \textit{rpsB-mCherry} (solid lines) strains.
As can be seen, fluorescence levels of the modified strains are far above the autofluorescence measured on the WT strain.
\textit{(C)}~The ratio of fluorescence over absorbance, a proxy for the fluorescence concentration in the cells.
Autofluorescence background was corrected by removing the autofluorescence measured on the WT wells.
We observe a strange increase in fluorescence concentration for the \textit{rpsB-mCherry} strain at stationary phase, something does not occur on the \textit{rpsB-gfp} strain.
Before each measurement cycle, the following procedure was applied: shaking (Orbital 6mm) for 30~s, shaking (Linear 6mm) for 30~s, waiting for 5~s.
}
\label{fig:suppinfo_caract}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/cell1}
\caption{
\textbf{Complete analysis for the cell located at the bottom of the well labelled "X1Y1, W11".}
(Figure description available in the introduction of \ref{S8_Text}.)\newline
Growth rate and resource allocation are particularly unstable at the beginning of the experiment, but seems to have stabilized before the upshift.
Oscillations in the RFU/pixel/cell signal are clearly visible, and result in oscillations in the resource allocation signal reconstruction.
}
\label{fig:cell1}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/cell2}
\caption{
\textbf{Complete analysis for the cell located at the bottom of the well labelled "X2Y1, W7".}
(Figure description available in the introduction of \ref{S8_Text}.)\newline
Here, the resource allocation is unstable at the beginning of the experiment despite the apparent regularity of the RFU/pixel/cell.
This could indicate a smoothing factor that is too low for this particular cell.
}
\label{fig:cell2}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/cell3}
\caption{
\textbf{Complete analysis for the cell located at the bottom of the well labelled "X3Y2, W5".}
(Figure description available in the introduction of \ref{S8_Text}.)\newline
We see for this cell that the reconstruction of the growth rate on the acetate medium (before 0) is affected by the huge gap in the acquisition, the increase before -150~min clearly being an artifact.
However, because the Kalman smoothing procedure we used allows for flexibility between the mother and the daughter cells, we quickly recover a more realistic growth rate before the upshift.
Despite an instability at the beginning of the experiment, the resource allocation reconstruction exhibits a remarkable stability during growth on the acetate medium, followed by oscillations after the upshift on glucose (also visible directly on the RFU/pixel/cell data).
}
\label{fig:cell3}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/cell4}
\caption{
\textbf{Complete analysis for the cell located at the bottom of the well labelled "X2Y2, W3".}
(Figure description available in the introduction of \ref{S8_Text}.)\newline
The reconstruction on this cell is a good instance of what was expected: the growth rate is stable on acetate, than quickly increases after the upshift on glucose.
The resource allocation is stable on acetate (indicating a steady state) than starts to oscillate after the upshift on glucose, even though the smoothing factor seems a little too high for this particular cell (some oscillations are poorly predicted after the upshift, in particular in the interval [200,400]~min).
}
\label{fig:cell4}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/cell5}
\caption{
\textbf{Complete analysis for the cell located at the bottom of the well labelled "X1Y2, W6".}
(Figure description available in the introduction of \ref{S8_Text}.)\newline
The results on this cell are particularly terrible: we do not seem to get a good steady state on acetate since the reconstructed resource allocation appears either negative or oscillating before the upshift.
It is thus impossible to conclude if the oscillations observed after the upshift are real or due to overfitting.
This cell highlights why robust steady states before and after the transition are important to conclude on the signal reconstruction.
}
\label{fig:cell5}
\end{figure}

>>>>

\chapter{Dynamics of the gene expression machinery during an acetate-glucose upshift}
\chaptermark{Measurements during an upshift}
\label{chap:experiments}

@@

Finalement, nous appliquons cette méthode à la reconstruction du taux de croissance puis de $\alpha (\cdot)$, et présentons les résultats dans la Section~\ref{sec:estim_gr_ra}.
En moyenne, le taux de croissance présente le comportement attendu (Figs~\ref{fig:growth_rate_estimation} et \ref{fig:growth_rate_estimation_median}), avec des valeurs à l'état de croissance stationnaire qui sont proches de celles connues pour \textit{E.~coli}, et une augmentation rapide au moment de la transition vers le milieu riche.
Les résultats pour la reconstruction de $\alpha (\cdot)$ sont cependant plus mitigés.
Même si un comportement oscillatoire est observé pour chaque cellule (Figs~\ref{fig:gene_activity} et \ref{fig:gene_activity_heatmap}), il n'est pas certain que ce dernier ne soit pas un artefact dû à une mauvaise calibration de la méthode.
Il aurait fallu en effet obtenir de longs états stationnaires de croissance avant et après la transition pour pouvoir réaliser cette calibration dans de bonnes conditions.
Néanmoins, si l'on s'intéresse au comportement médian des cellules (Fig.~\ref{fig:gene_activity_median}), on constate que la première oscillation survenant après la transition semble conservée dans la population (Fig.~\ref{fig:gene_activity_heatmap}).
Ces résultats s'avèrent finalement encourageants, et incitent à pousser plus en avant cette étude, notamment via l'acquisition de nouvelles données pour lesquelles les problèmes cités auront pu être corrigés (voir discussion en Section~\ref{sec:chap3_discussion}).
\end{chapter_summary}
\selectlanguage{english}

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/data_acquisition}
\caption{\small
\textbf{Results of data acquisition.}
We imaged 6 fields (X[1-3]Y[1-2]) each containing 15 wells (W[0-14]) for $\sim$40~h.
From the 90 wells, 68 were suitable for further analysis (the others being empty, out of frame for some period of time, or plugged).
For all wells, data points are missing in the interval between [-720,-150] because the camera was out of focus.
We stopped the analysis at 550~min, when about 3/4 of the bacteria were still growing, before the entire population died within a few hours for an unknown reason.
The image labeled "Raw image" is the last image analyzed for the highlighted well.
The bacterium on the left of this image was manually segmented by selecting two pixels at the poles on the fluorescence images (red cross).
A 6-pixel-wide rectangular mask was computed for each image, resulting in the "Segmented image" on the right.
Fluorescence intensities are expressed in Relative Fluorescence Units (RFU) on a 16-bit image and were corrected for camera background, but not autofluorescence background (Material and Methods~\ref{sec:cell_segmentation} and Discussion~\ref{sec:chap3_discussion}).
The fluorescence intensity of the cell, expressed in units RFU/pixel/cell, was computed by dividing the sum of the fluorescence intensities of the pixels in the rectangle by the total number of pixels in the rectangle.
The cell length is the distance in pixels between the two poles (red line).
The thick lines in green and black highlight the time-varying length and fluorescence density for the cell visible in the top images, labeled "X3Y2, W2".
The vertical dashed lines represent the time of the upshift from growth on acetate to growth on glucose.
}
\label{fig:data_acquisition}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/synthetic_upshift}
\caption{\small
\textbf{Performance of the Kalman smoothing procedure on synthetic data simulating an acetate-glucose upshift.}
\textit{(A)}~Synthetic data simulating an upshift from acetate to glucose, with and without additive white noise, as well as the results obtained by Kalman smoothing.
The synthetic data were generated by simulating the model presented in Eqs~\ref{eq:pdef}-\ref{eq:rdef} with the on-off regulatory strategy (Eq.~\ref{eq:stratswitch} and Fig.~\ref{fig:onoffresults}).
The model parameters used for the simulation are $e_{M,\texttt{Ace}} = 0.18$~h\textsuperscript{-1}, $e_{M,\texttt{Glu}} = 0.9$~h\textsuperscript{-1}, $k_R = 3.6$~h\textsuperscript{-1}, $\beta = 0.003$~L~g\textsuperscript{-1}, $K_R = 1$~g~L\textsuperscript{-1}.
The predicted $r(t)$ profile was multiplied by a factor $\gamma = 0.02$~RFU~L~g\textsuperscript{-1} in order to obtain the corresponding fluorescence intensity profile $F$ (dashed black curve).
The noise level was estimated from the data (\ref{S7_Text}) and added to $F$.
The choice of the parameters of the Kalman smoothing procedure is discussed in the Material and Methods~\ref{sec:meth_kalman}.
\textit{(B)}~Estimation of the resource allocation profile $\gamma \alpha / \beta$ based on the data in \textit{(A)}.
Following Chapter~\ref{chap:theory}, $\alpha (t)$ displays a bang-bang-singular profile during the upshift (dashed black curve).
While the Kalman smoother is not able to capture the discontinuous variations in $\gamma \alpha / \beta$, it qualitatively reproduces the input quite well (red solid curve).
\textit{(C)}~The predicted growth rate during the upshift experiment.
This information is used as an input in the smoothing procedure, since it is supposed to have been independently estimated from the measurements $\left\{L(t_0), ..., L(t_{N-1}) \right\}$.
}
\label{fig:synthetic_upshift}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[scale=0.985]{./Fig/suppinfo_caract}
\caption{\small
\textbf{Growth curves in M9~0.2\% glucose for the \textit{rpsB-gfp} and \textit{rpsB-mCherry} strains.}
Growth and measurements were performed at 37$^\circ$C on a 96-well microplate in a Tecan infinite 200 pro reader.
We monitored 4 wells for each of the WT, \textit{rpsB-gfp}, and \textit{rpsB-mCherry} strains.
Curves were time shifted to correct for variability in the inoculation process (-50 min for WT, and -20 min for \textit{rpsB-mCherry}).
\textit{(A)}~600~nm absorbance as measured by the Tecan microplate reader, corrected for background by removing M9 absorbance.
Curve superposition indicates similar growth rates between the WT and the modified strains.
\textit{(B)}~Fluorescence measured in each well, expressed in Relative Fluorescence Units (RFU).
Green fluorescence (485~nm excitation, 535~nm emission) is measured for the WT (dashed lines) and the \textit{rpsB-gfp} (solid lines) strains.
Red fluorescence (560~nm exc., 635~nm em.) is measured for the WT (dashed lines) and the \textit{rpsB-mCherry} (solid lines) strains.
Fluorescence levels of the modified strains are far above the autofluorescences measured on the WT strain.
\textit{(C)}~Ratio of fluorescence over absorbance, a proxy for the fluorescence concentration in the cells.
Autofluorescence background was corrected by removing the fluorescence measured on the WT.
We observe a strange increase in fluorescence concentration for the \textit{rpsB-mCherry} strain at stationary phase, something that does not occur on the \textit{rpsB-gfp} strain.
At each measurement cycle, the following procedure was applied: shaking (Orbital 6mm) for 30~s, shaking (Linear 6mm) for 30~s, waiting for 5~s.
}
\label{fig:suppinfo_caract}
\end{figure}

@@

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/cell1}
\caption{\small
\textbf{Complete analysis for the cell located at the bottom of the well labelled "X1Y1, W11".}
(Figure description available in the introduction of \ref{S8_Text}.)\newline
Growth rate and resource allocation are particularly unstable at the beginning of the experiment, but seems to have stabilized before the upshift.
Oscillations in the RFU/pixel/cell signal are clearly visible, and result in oscillations in the resource allocation signal reconstruction.
}
\label{fig:cell1}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/cell2}
\caption{\small
\textbf{Complete analysis for the cell located at the bottom of the well labelled "X2Y1, W7".}
(Figure description available in the introduction of \ref{S8_Text}.)\newline
Here, the resource allocation is unstable at the beginning of the experiment despite the apparent regularity of the RFU/pixel/cell.
This could indicate a smoothing factor that is too low for this particular cell.
}
\label{fig:cell2}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[scale=0.99]{./Fig/cell3}
\caption{\small
\textbf{Complete analysis for the cell located at the bottom of the well labelled "X3Y2, W5".}
(Figure description available in the introduction of \ref{S8_Text}.)\newline
We see for this cell that the reconstruction of the growth rate on the acetate medium (before 0) is affected by the huge gap in the acquisition, the increase before -150~min clearly being an artifact.
However, because the Kalman smoothing procedure we used allows for flexibility between the mother and the daughter cells, we quickly recover a more realistic growth rate before the upshift.
Despite an instability at the beginning of the experiment, the resource allocation reconstruction exhibits a remarkable stability during growth on the acetate medium, followed by oscillations after the upshift on glucose (also visible directly on the RFU/pixel/cell data).
}
\label{fig:cell3}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/cell4}
\caption{\small
\textbf{Complete analysis for the cell located at the bottom of the well labelled "X2Y2, W3".}
(Figure description available in the introduction of \ref{S8_Text}.)\newline
The reconstruction on this cell is a good instance of what was expected: the growth rate is stable on acetate, than quickly increases after the upshift on glucose.
The resource allocation is stable on acetate (indicating a steady state) than starts to oscillate after the upshift on glucose, even though the smoothing factor seems a little too high for this particular cell (some oscillations are poorly predicted after the upshift, in particular in the interval [200,400]~min).
}
\label{fig:cell4}
\end{figure}

\begin{figure}[p]
\centering
\includegraphics[scale=1]{./Fig/cell5}
\caption{\small
\textbf{Complete analysis for the cell located at the bottom of the well labelled "X1Y2, W6".}
(Figure description available in the introduction of \ref{S8_Text}.)\newline
The results on this cell are particularly terrible: we do not seem to get a good steady state on acetate since the reconstructed resource allocation appears either negative or oscillating before the upshift.
It is thus impossible to conclude if the oscillations observed after the upshift are real or due to overfitting.
This cell highlights why robust steady states before and after the transition are important to conclude on the signal reconstruction.
}
\label{fig:cell5}
\end{figure}
